---
title: "covid 19"
author: "Kelly Patrizio"
date: "October 20,2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: The Null Hypothesis is that there is no relationship between dying and the sex of the infected.  

$H_A$: The Alternative Hypothesis is that there is a relationship between dying and the sex of the infected.  

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata,main="frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display. 

```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The Bar Chart of Percent above indicates that a higher percentage of women lived that got COVID-19 in Korea and that a higher percentage of men died.  

## Numerical results

```{r}
table1 <- xtabs(~result + Sex,data=kRdata)
rowPerc(table1)
colPerc(table1)
```

In the cross-tab bar above we see that a higher percentage of women lived and a higher percentage of men died.  

Later we see that twice as many men died than women.   

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher"s Exact Test Above indicates that there is a 2.3 to 1 chance of dying.  If your a female than chance of dying is 25 out of 1831 and if your male the chance of dying is 44 out of 1411.  The chances of dying are small however, the men seem to have a greater chance of dying than the women. The p-value above would indicate that the ratio is not equal to 1 @ 9e-04.  

## Conclusions

Therefore we conclude that we reject the Null Hypothesis that there is no relationship between dying and the sex of the infected in favor of the Alternative, that there is a relationship between those who are dying and the infected.  We can also conclude that dying is not independent of sex. However, data shows that men are more likely to die from COVID-19 then women.   
    
 
